The limits of self commitment and private paternalism \thanks{}

### 2022-03-24 donde-2022-03-18-115935025.tex
restore comments from overleaf

\textcolor{red}{Figure} \ref{prob_ever_pawn} provides a visual take on the probability of re-pawning as a function of original treatment status.

@@

We now turn to the impact of the two arms that featured forms of `soft' commitment.  These results further reinforce the centrality of externally imposed commitment. 
Demand for the Soft Commitment contract  is higher than for the Forced Commitment contract by a factor of three: {31}\% of clients offered a choice between the status quo contract and the Soft Commitment contract chose commitment.  %Interestingly, the set of people who demand the Soft Commitment contract are different from those that demanded the Forced Commitment one. While in both cases the economically vulnerable demand less, in the case of the Soft Commitment those that we classify as present biased, those that report that they frequently fall into temptation buying, and those that ask for reminders of payment are \textit{less} likely to demand the Soft Commitment contract.  This suggests that those taking up the promise commitment are not the ones needing the commitment (see Figure \ref{interactions_takeup} in the Appendix). 
Nonetheless, we find Soft Commitment, whether chosen or induced, to be completely ineffective in this environment.  Table \ref{impact_soft} and  Figure \ref{effect_promises} finds that assigning clients to the choice-promise condition has a zero effect on financial cost, recovery, and the effective APR on average (Panel (a)), or on any sub-component of cost (Panel (b), and Tables \ref{table_robustness_fc_soft} and \ref{mechanisms_soft}). These results contrast sharply with the (lower-stakes, more short-term) evidence in laboratory experiments where promises have large effects on behavior. While clients are more enthusiastic about the Soft Commitment contract than the Forced Commitment one in the sense of tripling take-up, the promise contract provides no commitment. Results for the promise arms are consistent with \cite{Ted} and \cite{John} who find that people are likely to demand commitment when the commitment is too weak to have beneficial effects.  In short, not only is chosen hard commitment ineffective, but both forced and chosen soft commitment similarly fail to improve outcomes.  Only an externally imposed commitment product with reminders, imposed on people who would not have chosen it, decreases costs to borrowers.

@@

This is not the first paper to find low demand for commitment, however it is the first we know of that can compare choice vs forcing with a money-metric outcome such as loan repayment. The fact that the Forced Commitment contract lowers financial costs for almost all clients, together with the fact that only 11\% choose it must mean that many consumers are choosing contracts with higher financial costs for themselves. We now use several different techniques to shed further light on the idea that the `wrong' people are choosing commitment.

@@

More formally, let $Z_i$ be the observed, randomly assigned experimental allocation, where $Z_i=0$ denotes the control arm, $Z_i=1$ denotes the Forced arm, and $Z_i=2$ is the Choice arm.  Let $C_i$ be the choice type; $C_i=0$ when individual $i$ does not choose commitment and $C_i=1$ when chooses commitment, observed in the Choice arm and latent in the other two arms.  A fraction $p$ of individuals chooses the treatment in the Choice arm.  Finally, $D_i = Z_i\times \mathds{1}(Z_i\neq 2) + C_i\times \mathds{1}(Z_i=2)$ is the observed treatment indicator, and $Y(d,z)$ is the potential outcome function for $d=0,1$, and $z=0,1,2$.

@@

\[ToT = \frac{\beta^C}{p}; \qquad\quad TuT = \frac{(\beta^F - \beta^C)}{(1-p)}; \qquad\quad (ToT-TuT) = \frac{\beta^C}{p}- \frac{(\beta^F - \beta^C)}{(1-p)}. \]

Table \ref{tot_tut} shows the results of this analysis.  Given that the ITT results for the Choice arm imply an insignificant worsening of outcomes relative to the control, our estimates of the Treatment on the Treated in the first row of the bottom panel show negative results that are larger in absolute magnitude (multiplied by ~9, the inverse of the compliance rate) and similarly insignificant.  The Treatment on the Untreated (TuT), conversely, is strongly positive across the board, and indicates a significant improvement for all outcomes (here ITTs are effectively multiplied by 1.12, the inverse of the non-compliance rate).  Most novel is the ability to form a test of the difference between the ToT and the TuT, which is conducted in the bottom rows of the table.  We provide four different ways of estimating p-values for this comparison; the standard clustered estimates motivated by the experimental design, normal and percentile bootstraps, and randomization inference.  The differences between treatment effects for compliers and non-compliers are of borderline significance in some of the unadjusted specifications, but are not significant once adjusted.\footnote{It is worth noting that the statistical power of these estimands is driven by the compliance rate; given our low overall compliance we are better powered to measure the TuT than the ToT, and the power for the difference between the two would be highest with a compliance rate of 50\%.}   Because of the large standard errors on the ToT coming from the low compliance rate in the Choice arm we are unable to state with statistical confidence that the TuT is higher than the ToT, but the TuT is strongly positive while the ToT is weakly negative, and so in both relative and absolute terms the choices being made are `wrong' relative to the potential outcomes.

@@

Figure \ref{choose_wrong}(a) plots on the Y axis the fraction of clients that made a choice ``mistake'' and on the X axis the threshold $z$. When $z=0$, we calculate that about 82\% of clients in the Choice arm chose contracts that induced \textit{higher} financial cost. Most mistakes are failures to choose the fee-commitment contract ({97}\%), while only {4}\% of those who chose the fee-commitment contract would have been better in the status-quo one. The average mistake in pesos is \$250. %(90\% confidence interval [140,390]). 
As we increase the threshold on the size of the mistake necessary to be classified as having made the wrong choice, by definition the share in both groups falls, and among Choosers the number making an error falls quite quickly to zero.  However, the fraction of non-choosers making a mistake remains almost 20\% even at the highest threshold used (\$80 pesos), and panel b shows the \textcolor{red}{Complete this when understand figure}.  Panel C conducts a related but distinct exercise, showing the fraction of the entire control sample who would do better if assigned to the Forced arm; using the smallest threshold this fraction is over 95\%, showing that in this context there is an almost across-the board welfare benefit arising from this arm.  Since the only potential financial cost of the forced arm arises from individuals who would be induced to partially repay but then default anyways, this result suggests that very few such borrowers exist.

@@

We can represent the relationship between the HTEs from commitment and propensity scores in several ways.  Most straightforward is Figure \ref{HTE_vs_propensity}, which plots the individually-estimated treatment effect from Forced versus Control treatments on the Y-axis, and the predicted probability of choosing commitment on the x-axis.  The HTEs are very strongly downward-sloping over the part of the distribution that has heterogeneity in predicted propensities, the opposite of what we would expect from Roy-type selection.  The sample with the lowest propensity of choosing commitment has a smoothed benefit of roughly twice those with a .2 change of choosing, measured both in terms of benefit/cost and in terms of probability of repayment.  Hence the very large heterogeneous treatment effects are almost completely localized to the individuals who have the very lowest probability to select commitment.

@@

@@

\begin{figure}[H]
     \caption{Probability of ever pawning again}
    \label{prob_ever_pawn}
    \begin{center}
  \begin{subfigure}{0.49\textwidth}
     \caption{Reincidence after $n$ days}
        \centering
        \includegraphics[width=\textwidth]{Figuras/prob_reincidence_after.pdf}
    \end{subfigure}        
    \begin{subfigure}{0.49\textwidth}
     \caption{Reincidence within $n$ days}
        \centering
        \includegraphics[width=\textwidth]{Figuras/prob_reincidence_within.pdf}
    \end{subfigure}
    \end{center}
    \scriptsize
    This figure shows
      %\textit{Do file: }  \texttt{repeat\_loans.do}
\end{figure}

@@

\begin{figure}[H]
        \caption{Financial cost for different discount rates}
    \label{fc_discount_rates}
    \begin{center}
        \centering
        \includegraphics[width=0.55\textwidth]{Figuras/discount_effect.pdf}
    \end{center}
     \scriptsize This Figure estimates the treatment effect on financial cost with different discount rates.  
     %\textit{Do file: }  \texttt{discounted\_noeffect.do}
\end{figure}

@@

\newpage
\input{OA.tex}

>>>>

\textcolor{red}{Figure \ref{prob_ever_pawn} provides a visual take on the probability of re-pawning as a function of original treatment status.}

@@

We now turn to the impact of the two arms that featured forms of `soft' commitment.  These results further reinforce the centrality of externally imposed commitment. 
Demand for the Soft Commitment contract  is higher than for the Forced Commitment contract by a factor of 3: {31}\% of clients offered a choice between the status quo contract and the Soft Commitment contract chose commitment.  %Interestingly, the set of people who demand the Soft Commitment contract are different from those that demanded the Forced Commitment one. While in both cases the economically vulnerable demand less, in the case of the Soft Commitment those that we classify as present biased, those that report that they frequently fall into temptation buying, and those that ask for reminders of payment are \textit{less} likely to demand the Soft Commitment contract.  This suggests that those taking up the promise commitment are not the ones needing the commitment (see Figure \ref{interactions_takeup} in the Appendix). 
Nonetheless, we find Soft Commitment, whether chosen or induced, to be completely ineffective in this environment.  Table \ref{impact_soft} and  Figure \ref{effect_promises} finds that assigning clients to the choice-promise condition has a zero effect on financial cost, recovery, and the effective APR on average (Panel (a)), or on any sub-component of cost (Panel (b)). These results contrast sharply with the (lower-stakes, more short-term) evidence in laboratory experiments where promises have large effects on behavior. While clients are more enthusiastic about the Soft Commitment contract than the Forced Commitment one in the sense of tripling take-up, the promise contract provides no commitment. Results for the promise arms are consistent with \cite{Ted} and \cite{John} who find that people are likely to demand commitment when the commitment is too weak to have beneficial effects.  In short, not only is chosen hard commitment ineffective, but both forced and chosen soft commitment similarly fail to improve outcomes.  Only an externally imposed commitment product with reminders, imposed on people who would not have chosen it, decreases costs to borrowers.

@@

Our core impacts show that forced commitment is uniquely effective, a result that can only arise if those who would not have freely chosen commitment benefit from it.  We now delve more deeply into this result.

@@

More formally, let $Z_i$ be the observed, randomly assigned experimental allocation, where $Z_i=0$ denotes the control arm, $Z_i=1$ denotes the Forced arm, and $Z_i=2$ is the Choice arm.  Let $C_i$ be the choice type; $C_i=0$ when individual $i$ does not choose commitment and $C_i=1$ when chooses commitment, observed only in the Choice arm.  A fraction $p$ of individuals chooses the treatment in the Choice arm.  Finally, $D_i = Z_i\times \mathds{1}(Z_i\neq 2) + C_i\times \mathds{1}(Z_i=2)$ is the observed treatment indicator, and $Y(d,z)$ is the potential outcome function for $d=0,1$, and $z=0,1,2$.

@@

\[ToT = \frac{\beta^C}{p}; \hspace{15} TuT = \frac{(\beta^F - \beta^C)}{(1-p)}; \hspace{15} (ToT-TuT) = \frac{\beta^C}{p}- \frac{(\beta^F - \beta^C)}{(1-p)}. \]

Table \ref{tot_tut} shows the results of this analysis.  Given that the ITT results for the Choice arm imply an insignificant worsening of outcomes relative to the control, our estimates of the Treatment on the Treated in the first row of the bottom panel show negative results that are larger in absolute magnitude (multiplied by \~9, the inverse of the compliance rate) and similarly insignificant.  The Treatment on the Untreated (TuT), conversely, is strongly positive across the board, and indicates a significant improvement for all outcomes (here ITTs are multiplied by 1.12, the inverse of the non-compliance rate).  Most novel is the ability to form a test of the difference between the ToT and the TuT, which is conducted in the bottom rows of the table.  We provide four different ways of estimating p-values; the standard clustered estimates motivated by the experimental design, normal and percentile bootstraps, and randomization inference.  The differences between treatment effects for compliers and non-compliers are of borderline significance in some of the unadjusted specifications, but are not significant once adjusted.\footnote{It is worth noting that the statistical power of these estimands is driven by the compliance rate; given our low overall compliance we are better powered to measure the TuT than the ToT, and the power for the difference between the two would be highest with a compliance rate of 50\%.}

@@

\textcolor{red}{rewrite following paragraph with updated figure in hand.}
Figure \ref{choose_wrong}(a) plots on the left Y-axis the fraction of clients that made a choice ``mistake'' and on the right Y-axis its size in pesos. On the X-axis it varies the threshold $z$. When $z=0$, we calculate that about 75\% of clients in the Commitment Choice group chose contracts that induced \textit{higher} financial cost.  %(90\% confidence interval [45,94]). 
Most mistakes are failures to choose the fee-commitment contract ({85}\%), while only {15}\% of those who chose the fee-commitment contract would have been better in the status-quo one. The average mistake in pesos is \$250. %(90\% confidence interval [140,390]).

@@

We can represent the relationship between the HTEs from commitment and propensity scores in several ways.  Most straightforward is Figure \ref{HTE_vs_propensity}, which plots the individually-estimated treatment effect from Forced versus Control treatments on the Y-axis, and the predicted probability of choosing commitment on the X axis.  The HTEs are very strongly downward-sloping over the part of the distribution that has heterogeneity in predicted propensities, the opposite of what we would expect from Roy-type selection.  The sample with the lowest propensity of choosing commitment has a smoothed benefit of roughly twice those with a .2 change of choosing, measured both in terms of benefit/cost and in terms of probability of repayment.  Hence the very large heterogeneous treatment effects are almost completely localized to the individuals who have the very lowest probability to select commitment.

@@

\begin{table}[H]
\caption{Effects on several definitions of cost - Soft arms}
\label{table_robustness_fc_soft}
\begin{center}
\scriptsize{\input{./Tables/fc_robustness_soft.tex}}
\end{center}
 \scriptsize

%\textit{Do file: } \texttt{fc\_robustness\_soft.do}
\end{table}

\begin{table}[H]
\caption{Intermediate outcomes - Soft arms}
\label{mechanisms_soft}
\begin{center}
\scriptsize{\input{./Tables/mechanism_soft.tex}}
\end{center}

%\textit{Do file: } \texttt{mechanisms\_soft.do}
\end{table}

@@

\begin{figure}[H]
     \caption{Probability of ever pawning again}
    \label{prob_ever_pawn}
    \begin{center}
    \begin{subfigure}{0.49\textwidth}
     \caption{after taking first loan}
        \centering
        \includegraphics[width=\textwidth]{Figuras/prob_reincidence.pdf}
    \end{subfigure}
    \begin{subfigure}{0.49\textwidth}
     \caption{after recovery of first loan}
        \centering
        \includegraphics[width=\textwidth]{Figuras/prob_reincidence_des.pdf}
    \end{subfigure}    
    \end{center}
    \scriptsize
    This figure shows
      %\textit{Do file: }  \texttt{repeat\_loans.do}
\end{figure}

@@

@@

%\newpage

\input{OA.tex}
